---
title: "Main Report Analysis SonarCloud - Github"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an initial report as a test to analyze the causality of variables of development projects (GitHub) and quality characteristics of the software (Sonar Cloud). The data has been obtained through the public api of both platforms, and the json data has been filtered and pre-processed using an intermediate Mongo documentary database.

The result of the preprocessing has been stored in a CSV file. The first step is to import the data from that file.

```{r, echo=TRUE}
library(readr)
sonar_git <- read_delim("../data/sonar-git.csv", ";", quote = "\\\"", escape_double = FALSE, locale = locale(), trim_ws = TRUE)

```


Then, we filter our the matrix (with 107 variables) to manage those interesting (after some preliminary analyises were done). With the filtered data, we sown the descriptive statistics.

```{r sonar_git, echo=TRUE}
library(dplyr)
dataset1<-select(sonar_git, 
                  project,
                  version,
                  commits,
                  changes_by_commit,
                  committers,
                  committers_weight,
                  bugs,
                  code_smells,
                  complexity,
                  violations,
                  duplicated_blocks,
                  open_issues)


summary(dataset1)


```

## Including Plots

First we analyse commits/committers relationship


``` {r}
library(ggplot2)

library(plotly)

p <- ggplot(dataset1, aes(commits, project)) +
    geom_boxplot(fill="gray")
p

p <- ggplot(dataset1, aes(committers, project)) +
    geom_boxplot(fill="gray")
p



sp <- ggplot(dataset1, aes(x=commits, y=committers)) +
   geom_point(aes(colour=project)) +
   stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=0.2) + scale_fill_gradient(low="green", high="red")

sp + theme_classic()

zoom_sp <- sp + coord_cartesian(xlim = c(0, 120), ylim = c(0, 10))
zoom_sp + theme_classic()


kd <- with(dataset1, MASS::kde2d(committers, commits, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
 
fig


```


As preliminary analysis, we compute correlation values and draw a matrix of scatter plots:

```{r dataset1, echo=TRUE}

dataset_only_data<-select(dataset1, -1, -2)
M <- cor(dataset_only_data)
plot(dataset_only_data)

library(corrgram)
corrgram(dataset_only_data, order=FALSE, lower.panel=panel.shade,
   upper.panel=panel.pie, text.panel=panel.txt,
   main="correlation between variables")

```

```{r, echo=TRUE}
library(corrplot)
corrplot(M, method = "circle")
corrplot(M, method = "ellipse")
corrplot(M, method = "number")

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#AADD77", "#77AA44"))
res1 <- cor.mtest(dataset_only_data, conf.level = .95)
corrplot(M, method = "color", col = col(200),
          type = "upper", order = "original", number.cex = .8,
          addCoef.col = "black", # Add coefficient of correlation
          tl.col = "black", tl.srt = 90, # Text label color and rotation
          # Combine with significance
          p.mat = res1$p, sig.level = 0.05, insig = "blank", 
          # hide correlation coefficient on the principal diagonal
          diag = FALSE)



```


We focus on some variables where we observe certain correlation. First, we observe the behaviour of commits against complexity

``` {r}
library(ggplot2)
library(ggpubr)


monica<-dataset1[(dataset1[,'project']=='monica'), 1:12]
sp <- ggplot(monica, aes(x=commits, y=complexity)) + 
 geom_point(shape=16, aes(colour=project))+
 geom_smooth(method=lm,  linetype="dashed",
            color="darkred", fill="grey") +
coord_cartesian(xlim = c(0, 255), ylim = c(1500, 4300))
sp + stat_cor(method = "pearson", label.x = 160, label.y = 4300)


sonar_dotnet<-dataset1[(dataset1[,'project']=='sonar-dotnet'), 1:12]
sp <- ggplot(sonar_dotnet, aes(x=committers, y=duplicated_blocks)) + 
 geom_point(shape=16, aes(colour=project))+
 geom_smooth(method=lm,  linetype="dashed",
          color="darkred", fill="grey") +
 coord_cartesian(xlim = c(0, 15), ylim = c(35, 90))
sp + stat_cor(method = "pearson", label.x = 8, label.y = 75)

sonarqube<-dataset1[(dataset1[,'project']=='sonarqube'), 1:12]
sp <- ggplot(sonarqube, aes(x=committers, y=bugs)) + 
 geom_point(shape=16, aes(colour=project))+
 geom_smooth(method=lm,  linetype="dashed",
          color="darkred", fill="grey") +
coord_cartesian(xlim = c(0, 31), ylim = c(48, 64))
sp + stat_cor(method = "pearson", label.x = 12, label.y =58)


jacoco<-dataset1[(dataset1[,'project']=='jacoco'), 1:12]
sp <- ggplot(jacoco, aes(x=committers, y=code_smells)) + 
 geom_point(shape=16, aes(colour=project))+
 geom_smooth(method=lm,  linetype="dashed",
          color="darkred", fill="grey")
sp + stat_cor(method = "pearson", label.x = 3, label.y =230)



 ggplot(dataset1, aes(x=commits, y=complexity)) +
   geom_point(aes(colour=project))


 ggplot(dataset1, aes(x=commits, y=complexity, colour=project)) + 
   geom_point(shape=16)+
   geom_smooth(se = FALSE, method = lm)
 

sp <- ggplot(dataset1, aes(x=commits, y=complexity)) + 
   geom_point(shape=16, aes(colour=project))+
   geom_smooth(method=lm,  linetype="dashed",
              color="darkred", fill="grey")
sp

zoom_sp <- sp + coord_cartesian(xlim = c(0, 500), ylim = c(0, 20000))
zoom_sp


sp <- ggplot(dataset1, aes(x=commits, y=complexity)) +
   geom_point(aes(colour=project))

sp + geom_density_2d()
sp + stat_density_2d(aes(fill = ..level..), geom="polygon") + scale_fill_gradient(low="green", high="red")


committers_density <- ggplot(dataset1, aes(x=committers, fill=project)) + 
   geom_density(aes(group = project, 
                     colour = project, 
                     fill = project),
                     alpha=.1) + 
   theme(legend.position = "right")
committers_density

zoom_sp <- committers_density + coord_cartesian(xlim = c(0, 10), ylim = c(0, 0.75))
zoom_sp


commits_density <- ggplot(dataset1, aes(x=commits, fill=project)) + 
   geom_density(aes(group = project, 
                     colour = project, 
                     fill = project),
                     alpha=.1) + 
   theme(legend.position = "right")
commits_density

zoom_sp <- commits_density + coord_cartesian(xlim = c(0, 150), ylim = c(0, 0.025))
zoom_sp

```

We carry out a hierarchical clustering with all the variables and take 4 clusters


``` {r}

ddata1 <- dist(dataset_only_data)
gdata1 <- hclust(ddata1, method = "centroid")
plot(gdata1, sub = "example", xlab = "cases", ylab = "high") 
rect.hclust(tree = gdata1, k = 4,  border = c("red", "blue", "green", "orange"))
clusters <- cutree(tree = gdata1, k = 4)
clusters

dataset_clusters <- dataset_only_data
dataset_clusters$cluster <- factor(clusters)


sp <- ggplot(dataset_clusters, aes(x=commits, y=complexity)) + 
   geom_point(shape=16, aes(colour=cluster))+
   geom_smooth(method=lm,  linetype="dashed",
              color="darkred", fill="grey")

sp

```



##K-means scaled values
We carry out a K-means clustering with all the variables scaled and considering 4 clusters


``` {r}

library(cluster)

#Method for determine best number of clusters in K-means. Look for a bend or elbow in the sum of squared error (SSE) scree plot

mydata <- dataset_only_data
wss <- (nrow(mydata)-1)*sum(apply(mydata,2,var))
  for (i in 2:10) wss[i] <- sum(kmeans(mydata,
                                       centers=i)$withinss)
plot(1:10, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")



zdata1 <- scale(dataset_only_data)
kcdata1 <- kmeans(x = zdata1, centers = 4)
kcdata1$cluster
dataset_clusters$cluster2 <- factor(kcdata1$cluster)


sp <- ggplot(dataset_clusters, aes(x=commits, y=complexity, colour=cluster2, shape=cluster2)) + 
    geom_point(shape=16, aes(colour=cluster2))+
    geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="grey")

sp


```



We performed the characterization of clusters for the k-means algorithm


``` {r}

par(mfrow=c(1,1))
library(lattice)
splom(~ dataset_clusters[1:9], groups = cluster2, data = dataset_clusters, pch = 16)


library(vioplot)

x1 <- dataset_clusters$code_smells[dataset_clusters$cluster2==1]
x2 <- dataset_clusters$code_smells[dataset_clusters$cluster2==2]
x3 <- dataset_clusters$code_smells[dataset_clusters$cluster2==3]
x4 <- dataset_clusters$code_smells[dataset_clusters$cluster2==4]

vioplot(x1, x2, x3, x4, names=c("1", "2", "3", "4"), col="grey")

title("Code smells per cluster")


x1 <- dataset_clusters$commits[dataset_clusters$cluster2==1]
x2 <- dataset_clusters$commits[dataset_clusters$cluster2==2]
x3 <- dataset_clusters$commits[dataset_clusters$cluster2==3]
x4 <- dataset_clusters$commits[dataset_clusters$cluster2==4]
vioplot(x1, x2, x3, x4, names=c("1", "2", "3", "4"), col="grey")
title("Commits per cluster")


```

We compute correlation and scatter plots for clusters

``` {r, error=TRUE}

c1<-dataset_clusters[(dataset_clusters[,'cluster2']=='1'), 1:10]
c2<-dataset_clusters[(dataset_clusters[,'cluster2']=='2'), 1:10]
c3<-dataset_clusters[(dataset_clusters[,'cluster2']=='3'), 1:10]
c4<-dataset_clusters[(dataset_clusters[,'cluster2']=='4'), 1:10]

corrplot(cor(c1), method="number")
corrplot(cor(c2), method="number")
corrplot(cor(c3), method="number")
corrplot(cor(c4), method="number")


col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#AADD77", "#77AA44"))

res1 <- cor.mtest(c1, conf.level = .95)
corrplot(cor(c1), method = "color", col = col(200),
   type = "upper", order = "original", number.cex = .8,
   addCoef.col = "black",  tl.col = "black", tl.srt = 90,
   p.mat = res1$p, sig.level = 0.05, insig = "blank", diag = FALSE)

res1 <- cor.mtest(c2, conf.level = .95)
corrplot(cor(c2), method = "color", col = col(200),
   type = "upper", order = "original", number.cex = .8,
   addCoef.col = "black",  tl.col = "black", tl.srt = 90,
   p.mat = res1$p, sig.level = 0.05, insig = "blank", diag = FALSE)

res1 <- cor.mtest(c3, conf.level = .95)
corrplot(cor(c3), method = "color", col = col(200),
   type = "upper", order = "original", number.cex = .8,
   addCoef.col = "black",  tl.col = "black", tl.srt = 90,
   p.mat = res1$p, sig.level = 0.05, insig = "blank", diag = FALSE)

res1 <- cor.mtest(c4, conf.level = .95)
corrplot(cor(c4), method = "color", col = col(200),
   type = "upper", order = "original", number.cex = .8,
   addCoef.col = "black",  tl.col = "black", tl.srt = 90,
   p.mat = res1$p, sig.level = 0.05, insig = "blank", diag = FALSE)


sp <- ggplot(dataset_clusters, aes(x=commits, y=complexity, colour=cluster2, shape=cluster2)) + 
    geom_point()+
    geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="grey")
sp


```



Some 3d plots with correlations of several measures by cluster

``` {r, eval=FALSE, echo=FALSE}


c1<-dataset_clusters[(dataset_clusters[,'cluster2']=='1'), 1:10]
c2<-dataset_clusters[(dataset_clusters[,'cluster2']=='2'), 1:10]
c3<-dataset_clusters[(dataset_clusters[,'cluster2']=='3'), 1:10]
c4<-dataset_clusters[(dataset_clusters[,'cluster2']=='4'), 1:10]


kd <- with(c2, MASS::kde2d(committers, code_smells, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig

kd <- with(c3, MASS::kde2d(committers, code_smells, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig

kd <- with(c4, MASS::kde2d(committers, code_smells, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig



kd <- with(c1, MASS::kde2d(commits, complexity, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig

kd <- with(c2, MASS::kde2d(commits, complexity, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig

kd <- with(c3, MASS::kde2d(commits, complexity, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig

kd <- with(c4, MASS::kde2d(commits, complexity, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig


```


##Ploting for sonarqube project, cluster 2 and 3 differences

Ploting for sonarqube project, cluster 2 and 3 differences

``` {r, dataset_clusters}
library(ggplot2)
library(ggpubr)
theme_set(theme_minimal())

dataset_clusters

clusters<-dataset_clusters
clusters$project = dataset1$project
clusters$version = dataset1$version
sonar<-clusters[(clusters[,'project']=='sonarqube'), 1:14]


p_bugs<-ggplot(dat =sonar, aes(x=version, y=bugs)) + 
   geom_line(aes(group=1)) + 
   geom_point(aes(colour=cluster2, shape=cluster2, group=cluster2), size=3)
p_bugs

p_violations<-ggplot(dat =sonar, aes(x=version, y=violations)) + 
   geom_line(aes(group=1)) + 
   geom_point(aes(colour=cluster2, shape=cluster2, group=cluster2), size=3)
p_violations

p_commits<-ggplot(dat =sonar, aes(x=version, y=commits)) + 
   geom_line(aes(group=1)) + 
   geom_point(aes(colour=cluster2, shape=cluster2, group=cluster2), size=3)

p_commits

figure <- ggarrange(p_bugs, p_violations, p_commits, labels = c("a", "b", "c"), ncol = 1, nrow = 3)
figure


```

##k-means for normalized values

we perform the kmeans algorithm with normalized values and euclidean distance


``` {r}
library(vegan)

#data normalization
spe.norm <- decostand(dataset_only_data, "normalize") 
spe.ch <- vegdist(spe.norm, "euc")

spe.ch.ward <- hclust(spe.ch, method = "ward.D") 
plot(spe.ch.ward, sub = "Ward method")

#Calinski method
spe.KM.cascade <- cascadeKM(spe.norm, inf.gr = 2, sup.gr = 10, iter = 400, criterion = "ssi")
spe.KM.cascade$results
plot(spe.KM.cascade, sortg = TRUE)

#Silhouette plot
spe.kmeans <- kmeans(spe.norm, centers = 4, nstart = 100)
dissE <- daisy(spe.norm) 
sk <- silhouette(spe.kmeans$cl, dissE) 
plot(sk)




#compute k-means
set.seed(1) 
spe.kmeans <- kmeans(spe.norm, centers = 4, nstart = 100)

#clusters plot
spebc.ward.g <- cutree(spe.ch.ward,k = 4)
table(spe.kmeans$cluster, spebc.ward.g)
clusplot(spe.norm, spe.kmeans$cluster, color = TRUE, shade = TRUE, 
         labels = 2, lines = 0)


dataset_clusters$cluster3 <- factor(spe.kmeans$cluster)

```



We performed the characterization of clusters for the k-means algorithm


``` {r}

par(mfrow=c(1,1))
library(lattice)
splom(~ dataset_clusters[1:9], groups = cluster3, data = dataset_clusters, pch = 16)


library(vioplot)

x1 <- dataset_clusters$code_smells[dataset_clusters$cluster3==1]
x2 <- dataset_clusters$code_smells[dataset_clusters$cluster3==2]
x3 <- dataset_clusters$code_smells[dataset_clusters$cluster3==3]
x4 <- dataset_clusters$code_smells[dataset_clusters$cluster3==4]

vioplot(x1, x2, x3, x4, names=c("1", "2", "3", "4"), col="grey")

title("Code smells per cluster")


x1 <- dataset_clusters$commits[dataset_clusters$cluster3==1]
x2 <- dataset_clusters$commits[dataset_clusters$cluster3==2]
x3 <- dataset_clusters$commits[dataset_clusters$cluster3==3]
x4 <- dataset_clusters$commits[dataset_clusters$cluster3==4]
vioplot(x1, x2, x3, x4, names=c("1", "2", "3", "4"), col="grey")
title("Commits per cluster")


```

We compute correlation and scatter plots for clusters

``` {r, error=TRUE}

c1<-dataset_clusters[(dataset_clusters[,'cluster3']=='1'), 1:10]
c2<-dataset_clusters[(dataset_clusters[,'cluster3']=='2'), 1:10]
c3<-dataset_clusters[(dataset_clusters[,'cluster3']=='3'), 1:10]
c4<-dataset_clusters[(dataset_clusters[,'cluster3']=='4'), 1:10]

corrplot(cor(c1), method="number")
corrplot(cor(c2), method="number")
corrplot(cor(c3), method="number")
corrplot(cor(c4), method="number")


col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#AADD77", "#77AA44"))

res1 <- cor.mtest(c1, conf.level = .95)
corrplot(cor(c1), method = "color", col = col(200),
   type = "upper", order = "original", number.cex = .8,
   addCoef.col = "black",  tl.col = "black", tl.srt = 90,
   p.mat = res1$p, sig.level = 0.05, insig = "blank", diag = FALSE)

res1 <- cor.mtest(c2, conf.level = .95)
corrplot(cor(c2), method = "color", col = col(200),
   type = "upper", order = "original", number.cex = .8,
   addCoef.col = "black",  tl.col = "black", tl.srt = 90,
   p.mat = res1$p, sig.level = 0.05, insig = "blank", diag = FALSE)

res1 <- cor.mtest(c3, conf.level = .95)
corrplot(cor(c3), method = "color", col = col(200),
   type = "upper", order = "original", number.cex = .8,
   addCoef.col = "black",  tl.col = "black", tl.srt = 90,
   p.mat = res1$p, sig.level = 0.05, insig = "blank", diag = FALSE)

res1 <- cor.mtest(c4, conf.level = .95)
corrplot(cor(c4), method = "color", col = col(200),
   type = "upper", order = "original", number.cex = .8,
   addCoef.col = "black",  tl.col = "black", tl.srt = 90,
   p.mat = res1$p, sig.level = 0.05, insig = "blank", diag = FALSE)


sp <- ggplot(dataset_clusters, aes(x=commits, y=complexity, colour=cluster3, shape=cluster3)) + 
    geom_point()+
    geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="grey")
sp


```



Some 3d plots with correlations of several measures by cluster

``` {r, eval=FALSE, echo=FALSE}


c1<-dataset_clusters[(dataset_clusters[,'cluster3']=='1'), 1:10]
c2<-dataset_clusters[(dataset_clusters[,'cluster3']=='2'), 1:10]
c3<-dataset_clusters[(dataset_clusters[,'cluster3']=='3'), 1:10]
c4<-dataset_clusters[(dataset_clusters[,'cluster3']=='4'), 1:10]


kd <- with(c2, MASS::kde2d(committers, code_smells, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig

kd <- with(c3, MASS::kde2d(committers, code_smells, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig

kd <- with(c4, MASS::kde2d(committers, code_smells, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig



kd <- with(c1, MASS::kde2d(commits, complexity, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig

kd <- with(c2, MASS::kde2d(commits, complexity, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig

kd <- with(c3, MASS::kde2d(commits, complexity, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig

kd <- with(c4, MASS::kde2d(commits, complexity, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig


```
